### 2023-01-18 Poisot_2023_guidelines_prediction_draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@08949b36220530a13348d4076144bc29f375baf9 🚀

@@

\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[3] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc} % for \widthof, \maxof
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\maxof{\widthof{#1}}{\csllabelwidth}}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth}{#1}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}\geometry{verbose,letterpaper,tmargin=2.2cm,bmargin=2.2cm,lmargin=2.2cm,rmargin=2.2cm}

@@

@@

\begin{enumerate}
    \item The prediction of species interactions is gaining momentum as
a way to circumvent limitations in data volume. Yet, ecological networks
are challenging to predict because they are typically small and sparse.
Dealing with extreme class imbalance is a challenge for most binary
classifiers, and there are currently no guidelines as to how predictive
models can be trained for this specific problem.%
    \item Using simple mathematical arguments and numerical experiments
in which a variety of classifiers (for supervised learning) are trained
on simulated networks, we develop a series of guidelines related to the
choice of measures to use for model selection, and the ways to assemble
the training dataset.%
    \item Neither classifier accuracy nor the area under the receiver
operating characteristic curve (ROC-AUC) are informative measures for
the performance of interaction prediction. The area under the
precision-recall curve (PR-AUC) is a fairer assessment of performance.
In some cases, even standard measures can lead to selecting a more
biased classifier because the effect of connectance is strong. The
amount of correction to apply to the training dataset depends on network
connectance, on the measure to be optimized, and only weakly on the
classifier.%
    \item These results reveal that training machines to predict
networks is a challenging task, and that in virtually all cases, the
composition of the training set needs to be fine-tuned before performing
the actual training. We discuss these consequences in the context of the
low volume of data.%
\end{enumerate}

@@

An immense body of research on machine learning application to life
sciences is focused on genomics (which has very specific challenges, see
a recent discussion by Whalen et al., 2021); this sub-field has
generated recommendations that do not necessarily match the current
best-practices for SDMs, and therefore hint at the importance of
domain-specific guidelines. Chicco \& Jurman (2020) suggest using
Matthews correlation coefficient (MCC) over \(F_1\), as a protection
against over-inflation of predicted results; Delgado \& Tibau (2019)
advocate against the use of Cohen's \(\kappa\), again in favor of MCC,
as the relative nature of \(\kappa\) means that a worse classifier can
be picked over a better one; similarly, Boughorbel et al. (2017)
recommend MCC over other measures of performance for imbalanced data, as
it has more desirable statistical properties. More recently, Chicco et
al. (2021) temper the apparent supremacy of the MCC, by suggesting it
should be replaced by Youden's informedness (also known as \(J\),
bookmaker's accuracy, and the True-Skill Statistic) when the imbalance
in the dataset may not be representative of the actual imbalance. In a
way, the measures themselves need not be a strong focus for network
prediction, as they are routinely used in other field; the
discipline-specific question we seek to address is: `which metric should
be employed when predicting networks, and how to optimize it?'

@@

There is an immense diversity of measures to evaluate the performance of
classification tasks (Ferri et al., 2009). Here we will focus on five of
them with high relevance for imbalanced learning (He \& Ma, 2013). The
choice of metrics with relevance to class-imbalanced problems is
fundamental, because as Japkowicz (2013) unambiguously concluded,
``relatively robust procedures used for unskewed data can break down
miserably when the data is skewed.'' Following Japkowicz (2013), we
focus on two ranking metrics (the areas under the Receiver Operating
Characteristic and Precision Recall curves), and three threshold metrics
(\(\kappa\), informedness, and MCC; we will briefly discuss \(F_1\) but
show early on that it has undesirable properties).

@@

\begin{figure}
\hypertarget{fig:ecovalid}{%
\centering
\includegraphics{figures/valid_ensemble.png}
\caption{Visualisation of the raw (un-thresholded) models predictions
for one instance of a network prediction problem (shown in the
``Dataset'' panel). Increasing the value of the \(\xi\) parameter would
make the diagonal structure ``broader,'' leading to more interactions. A
visual inspection of the results is important, as it highlights how some
models can ``miss'' parts of the network; by combining them in an
ensemble, these gaps compensate one another, and lead (in this case) to
a better prediction.}\label{fig:ecovalid}
}
\end{figure}

The trained models were then thresholded (again by optimising
informedness), and their predictions transformed back into networks for
analysis; specifically, we measured the connectance, nestedness
(\(\eta\); Bastolla et al., 2009), modularity (\(Q\); Barber, 2007),
asymmetry (\(A\); Delmas et al., 2018), and Jaccard network
dissimilarity (Canard et al., 2014). This process was repeated 250
times, and the results are presented in tbl.~\ref{tbl:comparison}. The
k-NN model is an interesting instance here: it produces the network that
looks the most like the original dataset, despite having the lowest
PR-AUC, suggesting it hits high recall at the cost of low precision. The
ensemble was able to reach a very high PR-AUC (and a very high ROC-AUC),
which translated into more accurate reconstructions of the structure of
the network (with the exception of modulairty, which is underestimated
by \(0.03\)). This result bears elaborating. Measures of model
performance capture how much of the interactions and non-interactions
are correctly identified. As long as these predictions are not perfect,
some interactions will be predicted at the ``wrong'' position in the
network; these measures cannot describe the structural effect of these
mistakes. On the other hand, measures of network structure can have the
same value with interactions that fall at drastically different
positions; this is in part because a lot of these measures covary with
connectance, and in part because as long as these values are not 0 or
their respective maximum, there is a large number of network
configurations that can have the same value. That ROC-AUC is
consistently larger than PR-AUC may be a case of this measure masking
models that are not, individually, strong predictors (Jeni et al.,
2013). In this specif example, the combination of individually
``adequate'' models resulted in an extremely strong ensemble, suggesting
that the correct prediction of interactions (as measured by MCC, Inf.,
ROC-AUC, and PR-AUC) \emph{and} network properties is indeed a feasible
task under appropriately hyper-parameterized models.

\hypertarget{tbl:comparison}{}
\begin{longtable}[]{@{}rccccccccc@{}}
\caption{\label{tbl:comparison}Values of four performance metrics, and
five network structure metrics, for 500 independent predictions similar
to the ones presented in fig.~\ref{fig:ecovalid}. The values in
\textbf{bold} indicate the best value for each column (including ties).
Because the values have been rounded, values of 1.0 for the ROC-AUC
column indicate an average \(\ge 0.99\).}\tabularnewline
\toprule
Model & MCC & Inf. & ROC-AUC & PR-AUC & Conn. & \(\eta\) & \(Q\) & \(A\)
& Jaccard\tabularnewline
\midrule
\endfirsthead
\toprule
Model & MCC & Inf. & ROC-AUC & PR-AUC & Conn. & \(\eta\) & \(Q\) & \(A\)
& Jaccard\tabularnewline
\midrule
\endhead
Decision tree & 0.59 & 0.94 & 0.97 & 0.04 & 0.17 & 0.64 & 0.37 &
\textbf{0.42} & 0.1\tabularnewline
BRT & 0.46 & 0.91 & 0.97 & 0.36 & 0.2 & 0.78 & 0.29 & 0.41 &
0.19\tabularnewline
Random Forest & 0.72 & \textbf{0.98} & 0.99 & 0.1 & \textbf{0.16} &
\textbf{0.61} & 0.38 & \textbf{0.42} & \textbf{0.06}\tabularnewline
k-NN & 0.71 & \textbf{0.98} & 0.99 & 0.02 & \textbf{0.16} &
\textbf{0.61} & \textbf{0.39} & \textbf{0.42} &
\textbf{0.06}\tabularnewline
\emph{Ensemble} & \textbf{0.74} & \textbf{0.98} & \textbf{1.0} &
\textbf{0.79} & \textbf{0.16} & \textbf{0.61} & 0.38 & \textbf{0.42} &
\textbf{0.06}\tabularnewline
\emph{Data} & & & & & 0.16 & 0.56 & 0.41 & 0.42 & 0.0\tabularnewline
\bottomrule
\end{longtable}

@@

Finally, network connectance (\emph{i.e.} the empirical class imbalance)
should inform the composition of the training and testing set, because
it is an ecologically relevant value. In the approach outlined here, we
treat the class imbalance of the training set as an hyper-parameter, but
\emph{test} the model on a set that has the same class imbalance as the
actual dataset. This is an important distinction, as it ensure that the
prediction environment matches the testing environment (as we cannot
manipulate the connectance of the empirical dataset on which the
predictions will be made), and so the values measured on the testing set
(or validation set if the data volume allows one to exists) can be
directly compared to the values for the actual prediction. A striking
result from fig.~\ref{fig:optimbias} is that Informedness was almost
always maximal at 50/50 balance (regardless of connectance), whereas MCC
required \emph{more} positives to be maximized when connectance
\emph{increases}, matching the idea that it is a more stringent measure
of performance. This has an important consequence in ecological
networks, for which the pool of positive cases (interactions) to draw
from is typically small: the most parsimonious measure (\emph{i.e.} the
one requiring to discard the least amount of interactions to train the
model) will give the best validation potential, and in this light is
very likely informedness (maximizing informedness is, in fact, the
generally accepted default for imbalanced classification regardless of
the problem domain; Schisterman et al., 2005). This last result further
strengthens the assumption that the amount of bias \emph{is} an
hyper-parameter that must be fine-tuned, as using the wrong bias can
lead to models with lower performance; for this reason, it makes sense
to not train all models on the same training/testing set, but rather to
optimize the set composition for each of them.

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-Allouche2006AssAcc}{}%
Allouche, O., Tsoar, A., \& Kadmon, R. (2006). Assessing the accuracy of
species distribution models: Prevalence, kappa and the true skill
statistic (TSS). \emph{Journal of Applied Ecology}, \emph{43}(6),
1223--1232. \url{https://doi.org/10.1111/j.1365-2664.2006.01214.x}

\leavevmode\hypertarget{ref-Barber2007ModCom}{}%
Barber, M. J. (2007). Modularity and community detection in bipartite
networks. \emph{Physical Review E}, \emph{76}(6), 066102.
\url{https://doi.org/10.1103/PhysRevE.76.066102}

\leavevmode\hypertarget{ref-Bastolla2009ArcMut}{}%
Bastolla, U., Fortuna, M. A., Pascual-García, A., Ferrera, A., Luque,
B., \& Bascompte, J. (2009). The architecture of mutualistic networks
minimizes competition and increases biodiversity. \emph{Nature},
\emph{458}(7241), 1018--1020. \url{https://doi.org/10.1038/nature07950}

\leavevmode\hypertarget{ref-Beauchesne2016ThiOut}{}%
Beauchesne, D., Desjardins-Proulx, Archambault, P., \& Gravel, D.
(2016). Thinking Outside the Box--predicting Biotic Interactions in
Data-poor Environments. \emph{Vie Et Milieu-Life and enVironment},
\emph{66}(3-4), 333--342.

\leavevmode\hypertarget{ref-Becker2022OptPre}{}%
Becker, D. J., Albery, G. F., Sjodin, A. R., Poisot, T., Bergner, L. M.,
Chen, B., Cohen, L. E., Dallas, T. A., Eskew, E. A., Fagre, A. C.,
Farrell, M. J., Guth, S., Han, B. A., Simmons, N. B., Stock, M.,
Teeling, E. C., \& Carlson, C. J. (2022). Optimising predictive models
to prioritise viral discovery in zoonotic reservoirs. \emph{The Lancet
Microbe}. \url{https://doi.org/10.1016/S2666-5247(21)00245-7}

\leavevmode\hypertarget{ref-Bezanson2017JulFre}{}%
Bezanson, J., Edelman, A., Karpinski, S., \& Shah, V. (2017). Julia: A
Fresh Approach to Numerical Computing. \emph{SIAM Review}, \emph{59}(1),
65--98. \url{https://doi.org/10.1137/141000671}

\leavevmode\hypertarget{ref-Blaom2020MljJul}{}%
Blaom, A. D., Kiraly, F., Lienart, T., Simillides, Y., Arenas, D., \&
Vollmer, S. J. (2020). MLJ: A Julia package for composable machine
learning. \emph{Journal of Open Source Software}, \emph{5}(55), 2704.
\url{https://doi.org/10.21105/joss.02704}

\leavevmode\hypertarget{ref-Blaom2020FleMod}{}%
Blaom, A. D., \& Vollmer, S. J. (2020). \emph{Flexible model composition
in machine learning and its implementation in MLJ}.
\url{http://arxiv.org/abs/2012.15505}

\leavevmode\hypertarget{ref-Boughorbel2017OptCla}{}%
Boughorbel, S., Jarray, F., \& El-Anbari, M. (2017). Optimal classifier
for imbalanced data using Matthews Correlation Coefficient metric.
\emph{PloS One}, \emph{12}(6), e0177678.
\url{https://doi.org/10.1371/journal.pone.0177678}

\leavevmode\hypertarget{ref-Branco2015SurPre}{}%
Branco, P., Torgo, L., \& Ribeiro, R. (2015). \emph{A Survey of
Predictive Modelling under Imbalanced Distributions}.
\url{http://arxiv.org/abs/1505.01658}

\leavevmode\hypertarget{ref-Canard2014EmpEva}{}%
Canard, E. F., Mouquet, N., Mouillot, D., Stanko, M., Miklisova, D., \&
Gravel, D. (2014). Empirical evaluation of neutral interactions in
host-parasite networks. \emph{The American Naturalist}, \emph{183}(4),
468--479. \url{https://doi.org/10.1086/675363}

\leavevmode\hypertarget{ref-Caron2022AddElt}{}%
Caron, D., Maiorano, L., Thuiller, W., \& Pollock, L. J. (2022).
Addressing the Eltonian shortfall with trait-based interaction models.
\emph{Ecology Letters}, \emph{25}(4), 889--899.
\url{https://doi.org/10.1111/ele.13966}

\leavevmode\hypertarget{ref-Chicco2020AdvMat}{}%
Chicco, D., \& Jurman, G. (2020). The advantages of the Matthews
correlation coefficient (MCC) over F1 score and accuracy in binary
classification evaluation. \emph{BMC Genomics}, \emph{21}(1), 6.
\url{https://doi.org/10.1186/s12864-019-6413-7}

\leavevmode\hypertarget{ref-Chicco2021MatCor}{}%
Chicco, D., Tötsch, N., \& Jurman, G. (2021). The Matthews correlation
coefficient (MCC) is more reliable than balanced accuracy, bookmaker
informedness, and markedness in two-class confusion matrix evaluation.
\emph{BioData Mining}, \emph{14}, 13.
\url{https://doi.org/10.1186/s13040-021-00244-z}

\leavevmode\hypertarget{ref-deAguiar2019RevBia}{}%
de Aguiar, M. A. M., Newman, E. A., Pires, M. M., Yeakel, J. D.,
Boettiger, C., Burkle, L. A., Gravel, D., Guimarães, P. R., O'Donnell,
J. L., Poisot, T., Fortin, M.-J., \& Hembry, D. H. (2019). Revealing
biases in the sampling of ecological interaction networks. \emph{PeerJ},
\emph{7}, e7566. \url{https://doi.org/10.7717/peerj.7566}

\leavevmode\hypertarget{ref-Delgado2019WhyCoh}{}%
Delgado, R., \& Tibau, X.-A. (2019). Why Cohen's Kappa should be avoided
as performance measure in classification. \emph{PloS One}, \emph{14}(9),
e0222916. \url{https://doi.org/10.1371/journal.pone.0222916}

\leavevmode\hypertarget{ref-Delmas2018AnaEco}{}%
Delmas, E., Besson, M., Brice, M.-H., Burkle, L. A., Dalla Riva, G. V.,
Fortin, M.-J., Gravel, D., Guimarães, P. R., Hembry, D. H., Newman, E.
A., Olesen, J. M., Pires, M. M., Yeakel, J. D., \& Poisot, T. (2018).
Analysing ecological networks of species interactions. \emph{Biological
Reviews}, 112540. \url{https://doi.org/10.1111/brv.12433}

>>>>

\usepackage{gensymb}

@@

\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc} % for \widthof, \maxof
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\maxof{\widthof{#1}}{\csllabelwidth}}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth}{#1}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}\geometry{verbose,letterpaper,tmargin=2.2cm,bmargin=2.2cm,lmargin=2.2cm,rmargin=2.2cm}

@@

\begin{enumerate}
    \itemThe prediction of species interactions is gaining momentum as a
way to circumvent limitations in data volume. Yet, ecological networks
are challenging to predict because they are typically small and sparse.
Dealing with extreme class imbalance is a challenge for most binary
classifiers, and there are currently no guidelines as to how predictive
models can be trained for this specific problem.%
    \itemUsing simple mathematical arguments and numerical experiments
in which a variety of classifiers (for supervised learning) are trained
on simulated networks, we develop a series of guidelines related to the
choice of measures to use for model selection, and the ways to assemble
the training dataset.%
    \itemNeither classifier accuracy nor the area under the receiver
operating characteristic curve (ROC-AUC) are informative measures for
the performance of interaction prediction. The area under the
precision-recall curve (PR-AUC) is a fairer assessment of performance.
In some cases, even standard measures can lead to selecting a more
biased classifier because the effect of connectance is strong. The
amount of correction to apply to the training dataset depends on network
connectance, on the measure to be optimized, and only weakly on the
classifier.%
    \itemThese results reveal that training machines to predict networks
is a challenging task, and that in virtually all cases, the composition
of the training set needs to be fine-tuned before performing the actual
training. We discuss these consequences in the context of the low volume
of data.%
\end{enumerate}

@@

@@

An immense body of research on machine learning application to life
sciences is focused on genomics (which has very specific challenges, see
a recent discussion by Whalen et al., 2021); this sub-field has
generated recommendations that do not necessarily match the current
best-practices for SDMs, and therefore hint at the importance of
domain-specific guidelines. Chicco \& Jurman (2020) suggest using
Matthews correlation coefficient (MCC) over \(F_1\), as a protection
against over-inflation of predicted results; Delgado \& Tibau (2019)
advocate against the use of Cohen's \(\kappa\), again in favor of MCC,
as the relative nature of \(\kappa\) means that a worse classifier can
be picked over a better one; similarly, Boughorbel et al. (2017)
recommend MCC over other measures of performance for imbalanced data, as
it has more desirable statistical properties. More recently, Chicco et
al. (2021) temper the apparent supremacy of the MCC, by suggesting it
should be replaced by Youden's informedness (also known as \(J\),
bookmaker's accuracy, and the True-Skill Statistic) when the imbalance
in the dataset may not be representative of the actual imbalance. In a
way, the measures themselves need not be a strong focus for network
prediction, as they are routinely used in other field; the
discipline-specific question we seek to address is: `which metric should
be employed when predicting networks, and how to optimize it?'.

@@

There is an immense diversity of measures to evaluate the performance of
classification tasks (Ferri et al., 2009). Here we will focus on five of
them with high relevance for imbalanced learning (He \& Ma, 2013). The
choice of metrics with relevance to class-imbalanced problems is
fundamental, because as Japkowicz (2013) unambiguously concluded,
``relatively robust procedures used for unskewed data can break down
miserably when the data is skewed''. Following Japkowicz (2013), we
focus on two ranking metrics (the areas under the Receiver Operating
Characteristic and Precision Recall curves), and three threshold metrics
(\(\kappa\), informedness, and MCC; we will briefly discuss \(F_1\) but
show early on that it has undesirable properties).

@@

\begin{figure}
\hypertarget{fig:ecovalid}{%
\centering
\includegraphics{figures/valid_ensemble.png}
\caption{Visualisation of the raw (un-thresholded) models predictions
for one instance of a network prediction problem (shown in the
``Dataset'' panel). Increasing the value of the \(\xi\) parameter would
make the diagonal structure ``broader'', leading to more interactions. A
visual inspection of the results is important, as it highlights how some
models can ``miss'' parts of the network; by combining them in an
ensemble, these gaps compensate one another, and lead (in this case) to
a better prediction.}\label{fig:ecovalid}
}
\end{figure}

The trained models were then thresholded (again by optimising
informedness), and their predictions transformed back into networks for
analysis; specifically, we measured the connectance, nestedness
{[}\(\eta\); Bastolla et al. (2009){]}, modularity {[}\(Q\); Barber
(2007){]}, asymmetry {[}\(A\); Delmas et al. (2018){]}, and Jaccard
network dissimilarity (Canard et al., 2014). This process was repeated
250 times, and the results are presented in tbl.~\ref{tbl:comparison}.
The k-NN model is an interesting instance here: it produces the network
that looks the most like the original dataset, despite having the lowest
PR-AUC, suggesting it hits high recall at the cost of low precision. The
ensemble was able to reach a very high PR-AUC (and a very high ROC-AUC),
which translated into more accurate reconstructions of the structure of
the network (with the exception of modulairty, which is underestimated
by \(0.03\)). This result bears elaborating. Measures of model
performance capture how much of the interactions and non-interactions
are correctly identified. As long as these predictions are not perfect,
some interactions will be predicted at the ``wrong'' position in the
network; these measures cannot describe the structural effect of these
mistakes. On the other hand, measures of network structure can have the
same value with interactions that fall at drastically different
positions; this is in part because a lot of these measures covary with
connectance, and in part because as long as these values are not 0 or
their respective maximum, there is a large number of network
configurations that can have the same value. That ROC-AUC is
consistently larger than PR-AUC may be a case of this measure masking
models that are not, individually, strong predictors (Jeni et al.,
2013). In this specif example, the combination of individually
``adequate'' models resulted in an extremely strong ensemble, suggesting
that the correct prediction of interactions (as measured by MCC, Inf.,
ROC-AUC, and PR-AUC) \emph{and} network properties is indeed a feasible
task under appropriately hyper-parameterized models.

\hypertarget{tbl:comparison}{}
\begin{longtable}[]{@{}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.1548}}
  >{\centering\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0952}}
  >{\centering\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0952}}
  >{\centering\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0833}}
  >{\centering\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0952}}
  >{\centering\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0952}}
  >{\centering\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0952}}
  >{\centering\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0952}}
  >{\centering\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0952}}
  >{\centering\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0952}}@{}}
\caption{\label{tbl:comparison}Values of four performance metrics, and
five network structure metrics, for 500 independent predictions similar
to the ones presented in fig.~\ref{fig:ecovalid}. The values in
\textbf{bold} indicate the best value for each column (including ties).
Because the values have been rounded, values of 1.0 for the ROC-AUC
column indicate an average \(\ge 0.99\).}\tabularnewline
\toprule()
\begin{minipage}[b]{\linewidth}\raggedleft
Model
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
MCC
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Inf.
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
ROC-AUC
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
PR-AUC
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Conn.
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
\(\eta\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
\(Q\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
\(A\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Jaccard
\end{minipage} \\
\midrule()
\endfirsthead
\toprule()
\begin{minipage}[b]{\linewidth}\raggedleft
Model
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
MCC
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Inf.
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
ROC-AUC
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
PR-AUC
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Conn.
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
\(\eta\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
\(Q\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
\(A\)
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Jaccard
\end{minipage} \\
\midrule()
\endhead
Decision tree & 0.59 & 0.94 & 0.97 & 0.04 & 0.17 & 0.64 & 0.37 &
\textbf{0.42} & 0.1 \\
BRT & 0.46 & 0.91 & 0.97 & 0.36 & 0.2 & 0.78 & 0.29 & 0.41 & 0.19 \\
Random Forest & 0.72 & \textbf{0.98} & 0.99 & 0.1 & \textbf{0.16} &
\textbf{0.61} & 0.38 & \textbf{0.42} & \textbf{0.06} \\
k-NN & 0.71 & \textbf{0.98} & 0.99 & 0.02 & \textbf{0.16} &
\textbf{0.61} & \textbf{0.39} & \textbf{0.42} & \textbf{0.06} \\
\emph{Ensemble} & \textbf{0.74} & \textbf{0.98} & \textbf{1.0} &
\textbf{0.79} & \textbf{0.16} & \textbf{0.61} & 0.38 & \textbf{0.42} &
\textbf{0.06} \\
\emph{Data} & & & & & 0.16 & 0.56 & 0.41 & 0.42 & 0.0 \\
\bottomrule()
\end{longtable}

@@

Finally, network connectance (\emph{i.e.} the empirical class imbalance)
should inform the composition of the training and testing set, because
it is an ecologically relevant value. In the approach outlined here, we
treat the class imbalance of the training set as an hyper-parameter, but
\emph{test} the model on a set that has the same class imbalance as the
actual dataset. This is an important distinction, as it ensure that the
prediction environment matches the testing environment (as we cannot
manipulate the connectance of the empirical dataset on which the
predictions will be made), and so the values measured on the testing set
(or validation set if the data volume allows one to exists) can be
directly compared to the values for the actual prediction. A striking
result from fig.~\ref{fig:optimbias} is that Informedness was almost
always maximal at 50/50 balance (regardless of connectance), whereas MCC
required \emph{more} positives to be maximized when connectance
\emph{increases}, matching the idea that it is a more stringent measure
of performance. This has an important consequence in ecological
networks, for which the pool of positive cases (interactions) to draw
from is typically small: the most parsimonious measure (\emph{i.e.} the
one requiring to discard the least amount of interactions to train the
model) will give the best validation potential, and in this light is
very likely informedness {[}maximizing informedness is, in fact, the
generally accepted default for imbalanced classification regardless of
the problem domain; Schisterman et al. (2005){]}. This last result
further strengthens the assumption that the amount of bias \emph{is} an
hyper-parameter that must be fine-tuned, as using the wrong bias can
lead to models with lower performance; for this reason, it makes sense
to not train all models on the same training/testing set, but rather to
optimize the set composition for each of them.

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-Allouche2006AssAcc}{}}%
Allouche, O., Tsoar, A., \& Kadmon, R. (2006). Assessing the accuracy of
species distribution models: Prevalence, kappa and the true skill
statistic (TSS). \emph{Journal of Applied Ecology}, \emph{43}(6),
1223--1232. \url{https://doi.org/10.1111/j.1365-2664.2006.01214.x}

\leavevmode\vadjust pre{\hypertarget{ref-Barber2007ModCom}{}}%
Barber, M. J. (2007). Modularity and community detection in bipartite
networks. \emph{Physical Review E}, \emph{76}(6), 066102.
\url{https://doi.org/10.1103/PhysRevE.76.066102}

\leavevmode\vadjust pre{\hypertarget{ref-Bastolla2009ArcMut}{}}%
Bastolla, U., Fortuna, M. A., Pascual-García, A., Ferrera, A., Luque,
B., \& Bascompte, J. (2009). The architecture of mutualistic networks
minimizes competition and increases biodiversity. \emph{Nature},
\emph{458}(7241), 1018--1020. \url{https://doi.org/10.1038/nature07950}

\leavevmode\vadjust pre{\hypertarget{ref-Beauchesne2016ThiOut}{}}%
Beauchesne, D., Desjardins-Proulx, Archambault, P., \& Gravel, D.
(2016). Thinking Outside the Box--predicting Biotic Interactions in
Data-poor Environments. \emph{Vie Et Milieu-Life and enVironment},
\emph{66}(3-4), 333--342.

\leavevmode\vadjust pre{\hypertarget{ref-Becker2022OptPre}{}}%
Becker, D. J., Albery, G. F., Sjodin, A. R., Poisot, T., Bergner, L. M.,
Chen, B., Cohen, L. E., Dallas, T. A., Eskew, E. A., Fagre, A. C.,
Farrell, M. J., Guth, S., Han, B. A., Simmons, N. B., Stock, M.,
Teeling, E. C., \& Carlson, C. J. (2022). Optimising predictive models
to prioritise viral discovery in zoonotic reservoirs. \emph{The Lancet
Microbe}. \url{https://doi.org/10.1016/S2666-5247(21)00245-7}

\leavevmode\vadjust pre{\hypertarget{ref-Bezanson2017JulFre}{}}%
Bezanson, J., Edelman, A., Karpinski, S., \& Shah, V. (2017). Julia: A
Fresh Approach to Numerical Computing. \emph{SIAM Review}, \emph{59}(1),
65--98. \url{https://doi.org/10.1137/141000671}

\leavevmode\vadjust pre{\hypertarget{ref-Blaom2020MljJul}{}}%
Blaom, A. D., Kiraly, F., Lienart, T., Simillides, Y., Arenas, D., \&
Vollmer, S. J. (2020). MLJ: A Julia package for composable machine
learning. \emph{Journal of Open Source Software}, \emph{5}(55), 2704.
\url{https://doi.org/10.21105/joss.02704}

\leavevmode\vadjust pre{\hypertarget{ref-Blaom2020FleMod}{}}%
Blaom, A. D., \& Vollmer, S. J. (2020). \emph{Flexible model composition
in machine learning and its implementation in MLJ}.
\url{http://arxiv.org/abs/2012.15505}

\leavevmode\vadjust pre{\hypertarget{ref-Boughorbel2017OptCla}{}}%
Boughorbel, S., Jarray, F., \& El-Anbari, M. (2017). Optimal classifier
for imbalanced data using Matthews Correlation Coefficient metric.
\emph{PloS One}, \emph{12}(6), e0177678.
\url{https://doi.org/10.1371/journal.pone.0177678}

\leavevmode\vadjust pre{\hypertarget{ref-Branco2015SurPre}{}}%
Branco, P., Torgo, L., \& Ribeiro, R. (2015). \emph{A Survey of
Predictive Modelling under Imbalanced Distributions}.
\url{http://arxiv.org/abs/1505.01658}

\leavevmode\vadjust pre{\hypertarget{ref-Canard2014EmpEva}{}}%
Canard, E. F., Mouquet, N., Mouillot, D., Stanko, M., Miklisova, D., \&
Gravel, D. (2014). Empirical evaluation of neutral interactions in
host-parasite networks. \emph{The American Naturalist}, \emph{183}(4),
468--479. \url{https://doi.org/10.1086/675363}

\leavevmode\vadjust pre{\hypertarget{ref-Caron2022AddElt}{}}%
Caron, D., Maiorano, L., Thuiller, W., \& Pollock, L. J. (2022).
Addressing the Eltonian shortfall with trait-based interaction models.
\emph{Ecology Letters}, \emph{25}(4), 889--899.
\url{https://doi.org/10.1111/ele.13966}

\leavevmode\vadjust pre{\hypertarget{ref-Chicco2020AdvMat}{}}%
Chicco, D., \& Jurman, G. (2020). The advantages of the Matthews
correlation coefficient (MCC) over F1 score and accuracy in binary
classification evaluation. \emph{BMC Genomics}, \emph{21}(1), 6.
\url{https://doi.org/10.1186/s12864-019-6413-7}

\leavevmode\vadjust pre{\hypertarget{ref-Chicco2021MatCor}{}}%
Chicco, D., Tötsch, N., \& Jurman, G. (2021). The Matthews correlation
coefficient (MCC) is more reliable than balanced accuracy, bookmaker
informedness, and markedness in two-class confusion matrix evaluation.
\emph{BioData Mining}, \emph{14}, 13.
\url{https://doi.org/10.1186/s13040-021-00244-z}

\leavevmode\vadjust pre{\hypertarget{ref-deAguiar2019RevBia}{}}%
de Aguiar, M. A. M., Newman, E. A., Pires, M. M., Yeakel, J. D.,
Boettiger, C., Burkle, L. A., Gravel, D., Guimarães, P. R., O'Donnell,
J. L., Poisot, T., Fortin, M.-J., \& Hembry, D. H. (2019). Revealing
biases in the sampling of ecological interaction networks. \emph{PeerJ},
\emph{7}, e7566. \url{https://doi.org/10.7717/peerj.7566}

\leavevmode\vadjust pre{\hypertarget{ref-Delgado2019WhyCoh}{}}%
Delgado, R., \& Tibau, X.-A. (2019). Why Cohen's Kappa should be avoided
as performance measure in classification. \emph{PloS One}, \emph{14}(9),
e0222916. \url{https://doi.org/10.1371/journal.pone.0222916}

\leavevmode\vadjust pre{\hypertarget{ref-Delmas2018AnaEco}{}}%
Delmas, E., Besson, M., Brice, M.-H., Burkle, L. A., Dalla Riva, G. V.,
Fortin, M.-J., Gravel, D., Guimarães, P. R., Hembry, D. H., Newman, E.
A., Olesen, J. M., Pires, M. M., Yeakel, J. D., \& Poisot, T. (2018).
Analysing ecological networks of species interactions. \emph{Biological
Reviews}, 112540. \url{https://doi.org/10.1111/brv.12433}
